Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update results section

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/segment-example.png}
    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}.}
    \label{fig:text-segment}
\end{figure}

@@

A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text. A pooling operation was performed over time to find the most significant features in a region. After the convolutional layer, $m \times n_f$ are fed into bidirectional recurrent neural networks with 200 hidden units. Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not. We experimentally determined that the best performance is achieved when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting. Source codes of all experiments will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

@@

\section*{Results}
Our experimental results have two important directions. First, results are reported with respect to ``new segment'' and ``same segment'' classes as well as their weighted average in Table~\ref{tab:result_base}. Second, classification performance of different machine learning methods are outlined in Tables~\ref{tab:result_boundary}, \ref{tab:result_not_boundary} and \ref{tab:result_weighted_avg} when topic features are extracted with different probabilistic generative latent variable models.\\

@@

As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection. KNN also shows superior performance in all performance metrics for ``same segment'' class and weighted average over ``new segment'' and ``same segment'' classes. NB demonstrates the lowest performance among all models in terms of all performance metrics. On the other hand, SVM has the highest recall of 0.705 when only lexical features are used to identify ``new segment''. Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected since 96.71\% instances belong to ``same segment'' categories. For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in ``same segment'' identification compared to ``new segment'' detection. When lexical features are used in combination with punctuation mark, LSTM demonstrates the best performance in terms of precision and F1-measure for ``new segment'' identification while GRU shows 2.64\% higher recall than LSTM. GRU and LSTM show exactly same performance in terms of all performance metrics for ``same segement'' and overall classifications. \\

@@

Table~\ref{tab:result_weighted_avg} summarizes the weighted average results of the models for segmentation of e-Coaching emails. Overall, SVM has the best performance across all topic model-based features in terms of all performance metrics, achieves 0.990 precision, recall and F1-measure when DL-LDA-based topic features are used. Similar to results in Tables~\ref{tab:result_base}, \ref{tab:result_boundary} and \ref{tab:result_not_boundary}, NB demonstrates the lowest performance among all methods while best performance achieves with LCA-based topic features compared to other model-based topic features. Influence of the additional features is also consistent with the results in Tables~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}. Precision increases by 0.9\% and 0.2\%; recall increases by 0.9\% and 0.2\%; and F1-measure increases by 0.9\% and 0.3\% for SVM and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.

@@

Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods. Among all topic model-based feature, the DL-LDA-based topic feature provides the highest performance in ``new segment'' and ``same segment'' classification as well as weighted average over ``new segment'' and ``same segment'' classification results. In nearly all cases, ML methods perform better, when lexical features are used in combination with punctuation and topic model-based features. This results also indicate that segmentation performance might be improved by adding additional relevant features. \\

>>>>

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/segment-example.png}
    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}.}
    \label{fig:text-segment}
\end{figure}

@@

A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text. A pooling operation was performed over time to find the most significant features in a region. After the convolutional layer, $m \times n_f$ inputs are fed into bidirectional recurrent neural networks with 200 hidden units. Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not. We experimentally determined that the best performance is achieved when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting. Source codes of all experiments will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

@@

\section*{Results}
Our experimental results have two important directions. First, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}. Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.\\

@@

As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection. CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes. BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure. On the other hand, MLP has the highest precision of 0.836 when only word embeddings or lexical features are used to identify ``new segment''. CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes. Experimental results indicate that performance of classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified. For example, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN shows the lowest and CRF shows the second highest performance for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes.\\

@@

@@

Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods. Among all topic model-based feature, the DL-LDA-based topic feature provides the highest performance in ``new segment'' and ``same segment'' classification as well as weighted average over ``new segment'' and ``same segment'' classification results. In nearly all cases, ML methods perform better, when lexical features are used in combination with punctuation and topic model-based features. This results also indicate that segmentation performance might be improved by adding additional relevant features.
